---
title: "HW02"
author: "Zhijian Liu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. (Kutner et al, Problem 14.4)
#### a. 
Logistic mean response function:$E(Y_i)=\pi_i=F_L(\beta_0+\beta_1X_i)=\frac{exp(\beta_0+\beta_1X_i)}{1+exp(\beta_0+\beta_1X_i)}$

```{r 1.a, echo=F }
beta.0 <- -25
beta.1 <- 0.2
curve(exp(beta.0+beta.1*x)/(1+exp(beta.0+beta.1*x)), 50, 200, ylab = expression(pi))
```

#### b.
```{r 1.b}
logit <- function(p){log(p/(1-p))}
linear.component <- function(x){beta.0+beta.1*x}
GoFKernel::inverse(linear.component)(logit(0.5)) #inverse function
```

#### c.
```{r 1.c}
odds.1 <- exp(linear.component(150))
odds.2 <- exp(linear.component(151))
ratio.odds <- odds.2/odds.1
ratio.odds
exp(beta.1)
```

They are the same.

### 2. (Kutner et al, Problem 14.6)
#### a.
Probit mean response function:$E(Y_i)=\pi_i=\Phi(\beta_0+\beta_1X_i)$

```{r 2.a, echo=FALSE}
curve(pnorm(linear.component(x)), 100, 150, ylab = expression(pi),main = bquote(beta[0]~'='~-25~','~beta[1]~'='~0.2), col = 2)
curve(exp(beta.0+beta.1*x)/(1+exp(beta.0+beta.1*x)), col = 3, add = TRUE)
legend(100, 0.9, box.lty=0, legend = c('probit', 'logit'), col = c(2,3), lty =1)
```

The logistic curve is smoother and have a more constant marginal effect.

#### b.
```{r 2.b}
GoFKernel::inverse(linear.component)(GoFKernel::inverse(pnorm)(0.5))
#GoFKernel::inverse(linear.component)(qnorm(0.5)
```

### 3. (Kutner et al, Problem 14.7)
```{r 3, include=FALSE}
df <- read.table("/Users/liubeixi/Desktop/GLM/Data/GLM_sample_txt/CH14PR07.txt")
colnames(df) <- c("Y","X")
attach(df)
```

#### a.
```{r 3.a, echo = FALSE}
reg <- glm(Y~X, family = binomial(link=logit))
reg
```
Fitted response function: $\hat{\pi}=\frac{exp(-4.8+0.125X)}{1+exp(-4.8+0.125X)}$

#### b.
```{r 3.b, echo=FALSE}
loes <- loess(Y~X)
plot(X,Y,pch =20)
points(X, reg$fitted, col = 2, pch = 20)
points(X, loes$fitted, col =3, pch = 20)
legend(30, 0.8, box.lty=0, legend = c('observations', 'logistic', 'lowess'), col = c(1,2,3), pch = 20)
```

It appears to be a good fit.

#### c.
```{r 3.c}
data.frame("exp(b1)"=exp(coef(reg)[2]))
```
$e^{\beta_1}$ = 0.125. Every unit of increase in dues would cause the odds of not renewing the membership will change by a multiplier of 0.125.

#### d.
```{r 3.d, echo=F}
p <- exp(coef(reg)[1] + coef(reg)[2]*40)/(1+exp(coef(reg)[1] + coef(reg)[2]*40))
names(p) <- expression(pi)
p
```
If the dues are increased by $40, the estimated probability that association members will not renew their membership is 0.5487.

#### e.
```{r 3.e, echo=F}
x.hat <- (logit(0.75) - coef(reg)[1])/(coef(reg)[2])
names(x.hat) <- "X"
x.hat
detach(df)
```
The amount of dues increase for which 75 percent of the members are expected not to renew their association membership is 47.219.

### 4. (Kutner et al, Problem 14.12)
```{r 4, include=F}
df <- read.table("/Users/liubeixi/Desktop/GLM/Data/GLM_sample_txt/CH14PR12.txt")
colnames(df) <- c("X","n","Y")
attach(df)
```
#### a.
```{r 4.a, echo = F}
plot(X, Y/n)
```

The plot to some extent follows a S shape, so the logistic response function can be appropriate.

#### b.
```{r 4.b, echo=F}
reg <- glm(Y/n~X, family = binomial(link=logit), weight = n)
reg
```
Fitted response function: $\hat{mortality\;rate}=\frac{exp(-2.644+0.674X)}{1+exp(-2.644+0.674X)}$

#### c.
```{r 4.c, echo=F}
plot(X, Y/n)
points(X, reg$fitted, cex = 1.5, pch = 20)
legend(1, 0.8, box.lty=0, legend = c('proportion', 'logistic'), pch = c(1,20))
```

The fitted logistic response function appear to fit well.

#### d.
```{r 4.d, echo=F}
data.frame("exp(b1)"=exp(coef(reg)[2]))
```
$e^{\beta_1}$ = 1.96. Every unit of increase in the dose level would cause the odds of mortality will change by a multiplier of 1.96.

#### e.
```{r 4.e, echo=F}
p <- exp(coef(reg)[1] + coef(reg)[2]*3.5)/(1+exp(coef(reg)[1] + coef(reg)[2]*3.5))
names(p) <- expression(pi)
p
```
When the dose level is 3.5, the estimated probability that an insect dies is 0.429.

#### f.
```{r 4.f, echo=F}
x.hat <- (logit(0.5) - coef(reg)[1])/(coef(reg)[2])
names(x.hat) <- "X"
x.hat
detach(df)
```
The estimated median lethal dose is 3.922.

### 5. (Kutner et al, Problem 14.14)
```{r 5, include=F}
df <- read.table("/Users/liubeixi/Desktop/GLM/Data/GLM_sample_txt/CH14PR14.txt")
colnames(df) <- c("Y","X1","X2","X3")
attach(df)
```
#### a.
```{r 5.a, echo=F}
reg <- glm(Y~X1+X2+as.factor(X3), family = binomial(link=logit),data=df)
reg
```
Fitted response function: $\hat{\pi}=\frac{exp(-1.177+0.0728X_1-0.099X_2+0.434I_{X_3=1})}{1+exp(-1.177+0.0728X_1-0.099X_2+0.434I_{X_3=1})}$

#### b.
```{r 5.b, echo=F}
ef <- c(exp(coef(reg)[2]),exp(coef(reg)[3]),exp(coef(reg)[4]))
names(ef) <- c("exp(b1)","exp(b2)","exp(b3)")
ef
```

- Keeping other factors constant, every unit of increase in age would cause the odds of receiving a flu shor will change by a multiplier of 1.0755.  
- Keeping other factors constant, every unit of increase in health awareness would cause the odds of receiving a flu shor will change by a multiplier of 0.906.  
- Keeping other factors constant, for a male the odds of receiving a flu shor will higher than that for a female by 0.0755.

#### c.
```{r 5.c, echo=F}
p <- as.numeric(exp(coef(reg) %*% c(1,55,60,1))/(1+exp(coef(reg) %*% c(1,55,60,1))))
names(p) <- expression(pi)
p
```
The estimated probability that male clients aged 55 with a health awareness index of 60 will receive a flu shot is 0.064.
